---
title: 'Practical Machine Learning: Course Project'
author: "P.jirawat"
date: "January 22, 2019"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:\\Users\\Surface\\Desktop\\BOT\\Learning\\R\\8 Practical Machine Learning")
```

***

##Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the [website](http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har) (see the section on the Weight Lifting Exercise Dataset).

Data: [Training data](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv) and [Test data](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv)


The data for this project come from this [source](http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har).  
"Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu, R.; Fuks, H. Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements"

***

##Data Processing

###Libraries

```{r results = "hide",warning=FALSE,message=FALSE}
library(ggplot2)
library(caret)
library(rattle)
library(rpart)
library(randomForest)
library(gbm)
```

###Read data

```{r}
train.raw = read.csv(".\\Data\\pml-training.csv",header=TRUE)
valid.raw = read.csv(".\\Data\\pml-testing.csv",header=TRUE)
dim(train.raw)
```
```{r}
dim(valid.raw)
```

```{r}
str(train.raw)
```

###Clean Data
We will remove the variables that contains missing values and the first seven columns that do not contribute much to the  outcome.
```{r}
train.clean = train.raw[, colSums(is.na(train.raw)) == 0]
valid.dat = valid.raw[, colSums(is.na(valid.raw)) == 0]
train.clean = train.clean[, -c(1:7)]
valid.dat = valid.dat[, -c(1:7)]
dim(train.clean)
```
```{r}
dim(valid.dat)
```
Remove the variables that are near-zero-variance
```{r}
NZV = nearZeroVar(train.clean)
train.clean = train.clean[, -NZV]
dim(train.clean)
```

###Prepare Data
Split `train.clean` into train data (70%) and test data (30%). 
```{r}
set.seed(12345)
inTrain = createDataPartition(train.clean$classe, p = 0.7, list = FALSE)
train.dat = train.clean[inTrain, ]
test.dat = train.clean[-inTrain, ]
dim(train.dat)
```
```{r}
dim(test.dat)
```

###Correlation
Use findCorrelation function to search for highly correlated predictors
```{r}
cor = cor(train.dat[, -53])
h.cor = findCorrelation(cor, cutoff=0.75)
names(train.dat)[h.cor]
```

***

##Data Modeling

In the following sections, we will test 3 different models:  
- Classification tree  
- Random forest  
- Gradient boosting method  

###1. Train with Classification tree

Estimate
```{r}
control = trainControl(method="cv", 5)

model.ct = train(classe~., data=train.dat, method="rpart",trControl=control)
fancyRpartPlot(model.ct$finalModel)
```

Predict
```{r}
predict.ct = predict(model.ct, newdata=test.dat)
cm.ct = confusionMatrix(predict.ct, test.dat$classe)
cm.ct
```

Result
```{r}
re = data.frame(Method = c("CT","RF","GBM"), Accuracy = c(round(cm.ct$overall[[1]],3),"",""))
re
```

###2. Train with Random forest

Estimate
```{r}
model.rf = train(classe ~ ., data=train.dat, method="rf", trControl=control, verbose=FALSE)
model.rf
```

```{r}
model.rf$finalModel
```

```{r}
plot(model.rf)
```

Predict
```{r}
predict.rf = predict(model.rf, newdata=test.dat)
cm.rf = confusionMatrix(predict.rf, test.dat$classe)
cm.rf
```

Result
```{r}
re = data.frame(Method = c("CT","RF","GBM"), Accuracy = c(round(cm.ct$overall[[1]],3),round(cm.rf$overall[[1]],3),""))
re
```

###3. Train with Gradient boosting method

Estimte
```{r}
model.gbm  = train(classe ~ ., data=train.dat, method = "gbm", trControl = control, verbose=FALSE)
model.gbm
```

```{r}
model.gbm$finalModel
```

```{r}
plot(model.gbm)
```

Predict
```{r}
predict.gbm = predict(model.gbm, newdata=test.dat)
cm.gbm = confusionMatrix(predict.gbm, test.dat$classe)
cm.gbm
```

###Train Result
The result show that Random forest is the best method compared by the accuracy rate values.
```{r}
re = data.frame(Method = c("CT","RF","GBM"), Accuracy = c(round(cm.ct$overall[[1]],3),round(cm.rf$overall[[1]],3),round(cm.gbm$overall[[1]],3)))
re
```

***

##Conclusion
> Apply Random forest model to the validation data.

```{r}
re.final = predict(model.rf, newdata=valid.dat)
re.final
```
